Taking uncertainty in word embedding bias estimation seriously: a bayesian approach

### 2021-07-21 paperDraft/paperDraft.Rmd

We noticed that in @Manzini2019blackToCriminal words from all three sub-classes (in the case of religion by sub-class we mean Christianity, Judaism, Islam) were compared against all of the stereotypes, which means that there was no distinction between cases in which the stereotype is associated with a given sub-class, as opposed to the situation in which it is associated with another one. Not all of the stereotypical words have to be considered as harmful for all of the sub-groups. One should investigate the sub-groups separately as some of them may have stronger harmful associations than others.

@@

Our priors distributions for the Bayesian model reflect our knowledge about the data. The distribution for mean distance (cosine distance) is set to be between 0 and 2 with the mean equal to 1. We assume that most of the cosine distances may be considered neutral and therefore their similarity will be small. The distribution for prior for standard deviation is mostly between 0 and 1 with mean around 0. Finally the distribution for prior for coefficients is mostly between -1 and 1, with mean around 0. One should remember that the cosine distance values have range between 0 and 2 therefore the choice of the prior distributions should reflect this constraint. As we want to verify as objectively as possible the assumptions present in @Manzini2019blackToCriminal we decided to use fairly weak and slightly skeptical regularizing priors. The models parameters will have a posterior distribution obtained using either Hamiltionian Monte Carlo methods (STAN) available through the rethinking package.

>>>>

One thing to notice is that in @Manzini2019blackToCriminal words from all three subclasses (in the case of religion by a subclass we mean Christianity, Judaism, and Islam) were compared against all of the stereotypes, which means that there was no distinction between cases in which the stereotype is associated with a given sub-class, as opposed to the situation in which it is associated with another one. Not all of the stereotypical words have to be considered as harmful for all of the sub-groups. One should investigate the sub-groups separately as some of them may have stronger harmful associations than others.

@@

Our prior distributions for the Bayesian model reflect our first stab at what we know or expect before we see the data.  All our distributions are fairly weak. The distribution for mean distance (cosine distance) is set to fit (almost completely) between 0 and 2 with the mean equal to 1, in line with the idea that  most of the cosine distances may be considered neutral and therefore their similarity will be around 0. That is, we take it to be DISTRIBUTION. The  prior distribution for standard deviation is half-cauchy between 0 and 1, but unlikely to be as large as 1. Finally the prior distributions for  the coefficients have most of their mass between -1 and 1, with mean around 0.  As we want to put ourselves in skeptical shoes looking at   @Manzini2019blackToCriminal and asking whether the data would indeed convince us of the existence of bias, we decided to use slightly skeptical (mean 0), but  fairly weak (sd = ) as our regularizing priors. The model parameters will have a posterior distribution obtained using either Hamiltionian Monte Carlo methods (STAN) available through the rethinking package.

### 2021-07-21 paperDraft/paperDraft.tex

We noticed that in Manzini et al. (2019) words from all three
sub-classes (in the case of religion by sub-class we mean Christianity,
Judaism, Islam) were compared against all of the stereotypes, which
means that there was no distinction between cases in which the
stereotype is associated with a given sub-class, as opposed to the
situation in which it is associated with another one. Not all of the
stereotypical words have to be considered as harmful for all of the
sub-groups. One should investigate the sub-groups separately as some of
them may have stronger harmful associations than others.

@@

Our priors distributions for the Bayesian model reflect our knowledge
about the data. The distribution for mean distance (cosine distance) is
set to be between 0 and 2 with the mean equal to 1. We assume that most
of the cosine distances may be considered neutral and therefore their
similarity will be small. The distribution for prior for standard
deviation is mostly between 0 and 1 with mean around 0. Finally the
distribution for prior for coefficients is mostly between -1 and 1, with
mean around 0. One should remember that the cosine distance values have
range between 0 and 2 therefore the choice of the prior distributions
should reflect this constraint. As we want to verify as objectively as
possible the assumptions present in Manzini et al. (2019) we decided to
use fairly weak and slightly skeptical regularizing priors. The models
parameters will have a posterior distribution obtained using either
Hamiltionian Monte Carlo methods (STAN) available through the rethinking
package.

>>>>

One thing to notice is that in Manzini et al. (2019) words from all
three subclasses (in the case of religion by a subclass we mean
Christianity, Judaism, and Islam) were compared against all of the
stereotypes, which means that there was no distinction between cases in
which the stereotype is associated with a given sub-class, as opposed to
the situation in which it is associated with another one. Not all of the
stereotypical words have to be considered as harmful for all of the
sub-groups. One should investigate the sub-groups separately as some of
them may have stronger harmful associations than others.

@@

Our prior distributions for the Bayesian model reflect our first stab at
what we know or expect before we see the data. All our distributions are
fairly weak. The distribution for mean distance (cosine distance) is set
to fit (almost completely) between 0 and 2 with the mean equal to 1, in
line with the idea that most of the cosine distances may be considered
neutral and therefore their similarity will be around 0. That is, we
take it to be DISTRIBUTION. The prior distribution for standard
deviation is half-cauchy between 0 and 1, but unlikely to be as large as
1. Finally the prior distributions for the coefficients have most of
their mass between -1 and 1, with mean around 0. As we want to put
ourselves in skeptical shoes looking at Manzini et al. (2019) and asking
whether the data would indeed convince us of the existence of bias, we
decided to use slightly skeptical (mean 0), but fairly weak (sd = ) as
our regularizing priors. The model parameters will have a posterior
distribution obtained using either Hamiltionian Monte Carlo methods
(STAN) available through the rethinking package.
